\textbf{How open are hybrid journals included in transformative agreements?}

### 2024-09-18 analysis/manuscript.Rmd
Language check

The controversies surrounding hybrid open access and transformative agreements have led to varying policy conclusions. 
For instance, the UK's Joint Information Systems Committee (JISC) evaluated its open access strategy, which included transformative agreements [@Jisc_2024]. 
The evaluation revealed that while these agreements had a significant impact on the growth of open access in the UK, they had limited effects in facilitating a global shift towards full open access. 
As a result, the report advised that the UK's open access strategy should be reassessed.
After making similar observation, Norwegian and Swedish universities and their consortia also argued for policy changes [@norway; @Widding_2024]. 
Notably, the Association of Swedish Higher Education Institutions (Sveriges universitets- och högskoleförbund, SUHF) recommended only supporting agreements for publishing in fully open access journals by 2026 [@Widding_2024].
Furthermore, cOAlition S will conclude its financial support of transformative agreements at the end of 2024, but will continue to view open access resulting from such agreements as compliant. [^coalitions].
cOAlition S also removed the majority of hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
By contrast, the German DEAL consortium announced a five-year transformative agreement with Elsevier starting in 2024 and also renewed its contracts with Springer Nature and Wiley until the end of 2028. Similarly, the Colombia Consortium signed the first transformative agreements in Latin America [@Mu_oz_V_lez_2024].

Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the uptake of open access in hybrid journals, and the extent to which this can be attributed to transformative agreements.
Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @ludo_waltman_2022_7105355; @Taubert_2023; @Jisc_2024] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023; @Schmal_2024], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023]. 
In particular, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements. 
Likewise, article-level open access invoicing and cost data are only partly available, which would make it possible to establish a direct link between transformative agreements and open access publications [@Jahn_2021; @Kramer_2024].

@@

This study combines data from multiple publicly available sources, as shown in Figure \ref{fig:data_workflow}. 
Initially, transformative agreement data retrieved from cOAlition S Journal Checker Tool[^checkertool]  provided information about journal portfolios and participating institutions. 
After identifying hybrid journals by excluding fully open access journals, Crossref [@Hendricks_2020] served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
Because of a lack of comprehensive publicly available invoicing data, open access articles published through transformative agreements were determined by first author affiliations from OpenAlex [@priem2022openalex], linked to eligible institutions according to the transformative agreement data. 
In the following, the steps are described in detail.

@@

Data gathering started with obtaining journals included in transformative agreements from the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool [^checkertool], a voluntary effort based on publicly disclosed contracts in the ESAC Transformative Agreement Registry.[^esacregistry]. 
The dump consists of multiple online Google spreadsheets where each data file represents one agreement listed in the ESAC Transformative Agreement Registry.
From the retrieved spreadsheet files, journals and institutions involved per agreement were obtained.

It should be noted that although many library consortia see the need to register their agreements through the ESAC registry, many fail to disclose all details, including contracts [@Bakker_2024; @Kramer_2024]. 
Furthermore, the extent to which ESAC is comprehensive remains uncertain, potentially limiting the coverage of transformative agreements in the Journal Checker Tool. 
Another limitation of using the Journal Checker Tool and its underlying publicly available data dump to study the development of transformative agreements over time is that expired transformative agreements are constantly removed. 
To address this, four different snapshots were safeguarded and combined for this study: self-archived versions from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
This ensured that transformative agreements, which ended from 2021 onwards, were included, representing the majority of transformative agreements.
<!-- Agreements ending before 2021 could not be obtained.-->
Overall, the four combined Transformative Agreement Data dumps used in this study contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` of `r length(unique(esac_df$ID))` agreements listed in the ESAC Transformative Agreement Registry by December 2023.

@@

Affiliation metadata about corresponding authors are crucial for the planning and evaluation of transformative agreements because they are considered responsible for arranging open access publication [@Schimmer_2015; @Geschuhn_2017; @Borrego_2020]. 
For this study, country and institutional affiliations were retrieved from OpenAlex.
Because the corresponding authorship field was not fully supported by OpenAlex at the time of analysis, and respective affiliation data were only available for 54% of the investigated articles, this study focused on first authors and their affiliations instead; approximately 90% of the articles examined had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
First authors typically contribute the most to a paper and are often considered lead authors [@Larivi_re_2016], although notable differences in corresponding author metadata were observed across subjects, publication years and bibliometric databases [@Chinchilla_Rodr_guez_2024].
In case of lacking data about corresponding authors in bibliometric databases, related studies also utilised first authors as a proxy to examine open access payments and transformative agreements [@Zhang_2022; @Shu_2023; @Haucap_2021].

@@

To estimate the impact of transformative agreements on hybrid open access, participating institutions from the Transformative Agreement Data dump, which the cOAlition S crowd-sourced from  agreement documents and consortia, were matched with the first author affiliations recorded by OpenAlex using the ROR-ID.
Matching also considered the duration of agreements according to the ESAC registry. 
`r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` agreements were active between 2018 and 2022.
Upon inspection, Transformative Agreement Data did not comprehensively cover associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society.
To improve the matching, Transformative Agreement Data were automatically enriched with ROR-IDs from associated organisations according to OpenAlex's institution entity data.

@@

Between 2018 and 2022, `r n_hybrid_jns` out of `r n_hybrid_jns_all` eligible hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
These `r n_hybrid_jns` hybrid journals constituted the foundation for the subsequent analyses. 
Collectively, they provided open access to `r n_oa` out of `r n_total` articles during the investigated period, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` (`r n_ta_oa / n_oa * 100`%) open access articles.
Overall, this investigation was able to establish a link between open access articles and eligible institutions for `r n_tas_with_oa` out of `r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` (`r n_tas_with_oa / length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID)) * 100`%) active transformative agreements between 2018 and 2022 through the analysis of publication metadata, including affiliations and agreement data.

@@

Figure \ref{fig:results_overview}C depicts the substantial variations among the hybrid journals included in transformative agreements in terms of open access uptake.
Although the median generally follows the trend shown in Figure \ref{fig:results_overview}A, the farther stretch of upper quartiles and whiskers over the years illustrates that an increasing number of journals published an above-average proportion of open access articles. 
In 2022, 25% of hybrid journals (n = `r oa_jn_box_df |> filter(cr_year == "2022") |> mutate(top25 = quantile(prop, .75)) |> filter(prop >= top25) |> nrow()`) had an open access uptake of `r oa_jn_box_df |> filter(cr_year == "2022") |> pull(prop) |> quantile(0.75)*100`%, and `r (oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()) / length(unique(oa_jn_box_df$issn_l)) *100`% of journals (n = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()`) provided the majority of their articles under a Creative Commons license in the same year. 
On average, these journals were smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`). 
<!-- Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, and the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access). -->

@@

However, there are notable differences between the three large publishers. Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it recorded a comparably low number of open access articles, including those that are associated with transformative agreements.
In contrast, Springer Nature and Wiley provided open access to a larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open access), leading to higher open access market share (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp. `r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Wiley).
This difference between Elsevier on the one hand and Springer Nature and Wiley on the other can be attributed to transformative agreements, as the latter made the majority of their open access articles available through such deals (`r ((publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) * 100)`% Springer Nature  resp. `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) * 100`% Wiley).

@@

The varying degrees of adoption of open access across the three major publishers can be attributed to distinct approaches to transformative agreements. 
Springer Nature, for example, started offering open access agreements for its hybrid journal portfolio to selected consortia such as the Max Planck Society, the Swedish Bibsam Consortium and the Finnish FinELib consortium in 2015 under the name Springer Compact[^springercompact]. 
Prior to this, Springer had some pilot agreements with a small number of institutions, including the University of Göttingen [@Schmidt_2012].
However, the Springer Compact agreements were not included in the data as they concluded before the start of the transformative agreement data collection in June 2021. 
Nonetheless, the results demonstrate the importance of agreements for Springer Nature's hybrid open access business over the past five years (Figure \ref{fig:publisher_figure}).
In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access in Springer Nature hybrid journals was enabled through transformative agreements.
In the same year, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value)`) of Wiley's open access articles could be linked to transformative agreements. 
By contrast, Elsevier published a comparatively lower proportion of its open access articles through transformative agreements in 2022 (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) `; `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()*100`%).

@@

Figure \ref{fig:subject_panel} presents the relative growth of hybrid open access by subject area between 2018 and 2022. In particular, Social Sciences including Arts and Humanties journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`). 
<!-- In particular, journals in the subject areas recorded large high rate (see supplement)-->
Growth in Social Sciences category can be largely attributed to transformative agreements. 
In 2022, two-thirds of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) were published by the first authors affiliated with participating institutions (see \ref{fig:subject_panel}B).

```{r subject_panel, fig.asp=0.8, fig.cap="Development of open access in hybrid journals in transformative agreements between 2018 and 2022 by ASJC subject area. The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established. (A) Proportion of open access articles in hybrid journals per year and subject area (B) Percentage of hybrid open access via agreements per year and subject area Boxplots (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and subject area The individual outliers are not shown. Note that data on transformative agreements ending before June 2021 were not available for this study."}

@@

Between 2018 and 2022, high-income countries almost exclusively dominated hybrid open access publishing through transformative agreements. 
To discern socio-economic differences, these countries were grouped according to their membership in the Organisation for Economic Co-operation and Development (OECD).
During 2018 and 2022, first authors affiliated with institutions from OECD member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
This disparity between OECD nations and other countries becomes even more evident when considering open access through transformative agreements, as  `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa)` of `r n_ta_oa`, or `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa) / n_ta_oa*100`% of open access articles were associated with such agreements.

>>>>

The controversies surrounding hybrid open access and transformative agreements have led to varying policy conclusions. 
For instance, the British Joint Information Systems Committee (JISC) evaluated its open access strategy, which included transformative agreements [@Jisc_2024]. 
The evaluation revealed that while these agreements had a significant impact on the growth of open access in the country, they had limited effects in facilitating a global shift towards full open access. 
As a result, the report advised that the British open access strategy should be reassessed.
After making similar observation, Norwegian and Swedish universities and their consortia also argued for policy changes [@norway; @Widding_2024].
Furthermore, cOAlition S conclude its financial support of transformative agreements at the end of 2024, but continue to view open access resulting from such agreements as compliant. [^coalitions].
cOAlition S also removed the majority of hybrid journals from its Transformative Journal program in 2023 due to publishers' failure to meet self-defined open access growth targets [@Brainard_2023]. 
By contrast, the German DEAL consortium announced a five-year transformative agreement with Elsevier starting in 2024 and also renewed its contracts with Springer Nature and Wiley until the end of 2028. Similarly, the Colombia Consortium signed the first transformative agreements in Latin America [@Mu_oz_V_lez_2024].

Despite these controversies around transformative agreements as a means of transitioning journal publishing to full open access, there is limited evidence available on the uptake of open access in hybrid journals, and the extent to which this can be attributed to transformative agreements.
Previous studies have focused on specific countries [@P_l_nen_2020; @Huang_2020; @Haucap_2021; @Wenaas_2022; @ludo_waltman_2022_7105355; @Taubert_2023; @Jisc_2024] or publisher portfolios [@Pieper_2018; @Bakker_2024; @Fraser_2023; @Jahn_2021; @Momeni_2023; @Schmal_2024], while large-scale studies relied on self-reported agreement data [@Moskovkin_2022], or used APC pricing lists [@Shu_2023]. 
In particular, data availability is a limiting factor when studying the impact of transformative agreements [@Bakker_2024], because bibliometric databases, even though many allow the retrieval of open access articles in hybrid journals, do not directly attribute them to specific transformative agreements. 
Likewise, article-level open access invoicing and cost data, which would make it possible to establish a direct link between transformative agreements and open access publications [@Jahn_2021; @Kramer_2024].

@@

This study combines data from multiple publicly available sources, as shown in Figure \ref{fig:data_workflow}. 
Initially, transformative agreement data retrieved from cOAlition S Journal Checker Tool[^checkertool]  provided information about journal portfolios and participating institutions. 
After identifying hybrid journals by excluding fully open access journals, Crossref [@Hendricks_2020] served as the primary data source for article-level metadata including Creative Commons (CC) license information to indicate open access availability on publisher websites.
Because of a lack of comprehensive publicly available invoicing data, open access articles published through transformative agreements were determined by linking first author affiliations from OpenAlex [@priem2022openalex] to eligible institutions according to the transformative agreement data. 
In the following, the steps are described in detail.

@@

Data gathering started with obtaining journals included in transformative agreements from the publicly available Transformative Agreement Data dump[^tadump] used by the cOAlition S Journal Checker Tool[^checkertool], a voluntary effort based on publicly disclosed contracts in the ESAC Transformative Agreement Registry[^esacregistry]. 
The dump consists of multiple online Google spreadsheets where each data file represents one agreement listed in the ESAC Transformative Agreement Registry.
From the retrieved spreadsheet files, journals and institutions involved per agreement were obtained.

It should be noted that although many library consortia see the need to register their agreements through the ESAC registry, some fail to disclose all details, including the full-text of contracts [@Bakker_2024; @Kramer_2024]. 
Furthermore, the extent to which ESAC is comprehensive remains uncertain, potentially limiting the coverage of transformative agreements in the Journal Checker Tool. 
Another limitation of using the Journal Checker Tool and its underlying publicly available data dump to study the development of transformative agreements over time is that expired transformative agreements are constantly removed. 
To address this, four different snapshots were safeguarded and combined for this study: self-archived versions from July 2021, July 2022, and May 2023, as well as the most current dump downloaded on 11 December 2023.
This ensured that transformative agreements, which ended from 2021 onwards, were included, representing the majority of transformative agreements.
<!-- Agreements ending before 2021 could not be obtained.-->
Overall, the four combined Transformative Agreement Data dumps used in this study contained `r  length(unique(hoaddata::jct_hybrid_jns$esac_id))` of `r length(unique(esac_df$ID))` agreements listed in the ESAC Transformative Agreement Registry by December 2023.

@@

Affiliation metadata about corresponding authors are crucial for the planning and evaluation of transformative agreements because they are considered responsible for arranging open access publication [@Schimmer_2015; @Geschuhn_2017; @Borrego_2020]. 
For this study, country and institutional affiliations were retrieved from OpenAlex.
Because the corresponding authorship field was not fully supported by OpenAlex at the time of analysis, and respective affiliation data were only available for 54% of the investigated articles, this study focused on first authors and their affiliations instead; approximately 90% of the articles examined had first author affiliation metadata in OpenAlex, which is a much larger proportion than previously reported for the October 2022 snapshot [@Zhang_2024].
First authors typically contribute the most to a paper and are often considered as lead authors [@Larivi_re_2016; @Chinchilla_Rodr_guez_2024].
In case of lacking data about corresponding authors in bibliometric databases, related studies also utilised first authors as a proxy to examine open access payments and transformative agreements [@Zhang_2022; @Shu_2023; @Haucap_2021].

@@

To estimate the impact of transformative agreements on hybrid open access, participating institutions from the Transformative Agreement Data dump, which the cOAlition S crowd-sourced from  agreement documents and consortia, were matched with the first author affiliations recorded by OpenAlex using the ROR-ID.
Matching also considered the duration of agreements according to the ESAC registry. 
In total, `r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` agreements were active between 2018 and 2022.
Upon inspection, Transformative Agreement Data did not comprehensively cover associated institutions, such as university hospitals or institutes of large research organisations such as the Max Planck Society.
To improve the matching, Transformative Agreement Data were automatically enriched with ROR-IDs from associated organisations according to OpenAlex's institution entity data.

@@

Between 2018 and 2022, `r n_hybrid_jns` out of `r n_hybrid_jns_all` hybrid journals in transformative agreements published at least one open access article under a Creative Commons license.
These eligible `r n_hybrid_jns` hybrid journals constituted the foundation for the subsequent analyses. 
Collectively, they provided open access to `r n_oa` out of `r n_total` articles during the investigated period, representing a five-year open access proportion of `r n_oa / n_total * 100`%.
Authors who could make use of transformative agreements at the time of publication contributed `r n_ta_oa` (`r n_ta_oa / n_oa * 100`%) open access articles.
Overall, this investigation was able to establish a link between open access articles and eligible institutions for `r n_tas_with_oa` out of `r length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID))` (`r n_tas_with_oa / length(unique(tas_in_esac_and_hoad_active_in_2018_22$ID)) * 100`%) active transformative agreements between 2018 and 2022 through the linking of first author affiliations with agreement data.

@@

Figure \ref{fig:results_overview}C depicts the substantial variations among the hybrid journals included in transformative agreements in terms of open access uptake.
Although the median generally follows the trend shown in Figure \ref{fig:results_overview}A, the farther stretch of upper quartiles and whiskers over the years illustrates that an increasing number of journals published an above-average proportion of open access articles. 
In 2022, 25% of hybrid journals (n = `r oa_jn_box_df |> filter(cr_year == "2022") |> mutate(top25 = quantile(prop, .75)) |> filter(prop >= top25) |> nrow()`) had an open access uptake of `r oa_jn_box_df |> filter(cr_year == "2022") |> pull(prop) |> quantile(0.75)*100`%, and `r (oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()) / length(unique(oa_jn_box_df$issn_l)) *100`% of journals (n = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> nrow()`) provided the majority of their articles under a CC license in the same year. 
On average, these journals were smaller (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop >= 0.5) |> pull(n) |> sd()`) than those with an open access share below 50% (M = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> mean()`, SD = `r oa_jn_box_df |> filter(cr_year == "2022", prop <= 0.5) |> pull(n) |> sd()`). 
<!-- Notable exception of large journals with an above-average open access proportion were Physical Review D, a high-energy physics journal covered by the SCOAP3 consortium that provided open access to `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "2470-0010") |> pull(n)` articles in 2022, Astronomy and Physics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(oa_all)` out of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0004-6361") |> pull(n)` articles in 2022 were open access), which shifted to a subscribe to open business model for all accepted articles as of April 2022, and the Journal of Fluid Mechanics (`r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(oa_all)` of `r oa_jn_box_df |> filter(cr_year == "2022", issn_l == "0022-1120") |> pull(n)` articles in 2022 were open access). -->

@@

However, there are notable differences between the three large publishers. Although Elsevier published the largest volume of articles (n = `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(publisher_total)`, `r publisher_league_tbl  |> filter(publisher == "Elsevier") |> pull(cum)*100`%), it recorded a comparably low number of open access articles, including those that are associated with transformative agreements.
In contrast, Springer Nature and Wiley provided open access to a larger proportion of their articles (`r (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(publisher_total)) * 100`% of Springer Nature articles and `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(publisher_total)) * 100`% of Wiley articles were open access), leading to higher open access market shares (`r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Springer Nature resp. `r publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_perc)*100`% Wiley).
This difference between Elsevier on the one hand and Springer Nature and Wiley on the other can be attributed to transformative agreements, as the latter made the majority of their open access articles available through such deals (`r ((publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Springer Nature") |> pull(oa_total)) * 100)`% Springer Nature  resp. `r (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(ta_oa_total)) / (publisher_league_tbl  |> filter(publisher == "Wiley") |> pull(oa_total)) * 100`% Wiley).

@@

The varying degrees of adoption of open access across the three major publishers can be attributed to distinct approaches to transformative agreements. 
Springer Nature, for example, started offering open access agreements for its hybrid journal portfolio to selected consortia such as the Max Planck Society, the Swedish Bibsam consortium and the Finnish FinELib consortium in 2015 under the name Springer Compact[^springercompact]. 
Prior to this, Springer had some pilot agreements with a small number of institutions, including the University of Göttingen [@Schmidt_2012].
However, the Springer Compact agreements were not included in the data as they concluded before the start of the transformative agreement data collection in June 2021. 
Nonetheless, the results demonstrate the importance of agreements for Springer Nature's hybrid open access business over the five-years period (Figure \ref{fig:publisher_figure}B).
In 2022, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature") |> pull(value) |> sum()*100`% (n = `r  publisher_summary_df |> filter(cr_year == "2022", publisher == "Springer Nature", type == "ta_oa") |> pull(value)`) of open access in Springer Nature hybrid journals was enabled through transformative agreements.
In the same year, `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley") |> pull(value) |> sum()*100`% (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Wiley", type == "ta_oa") |> pull(value)`) of Wiley's open access articles could be linked to transformative agreements. 
By contrast, Elsevier published a comparatively lower proportion of its open access articles through transformative agreements in 2022 (n = `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) `; `r publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier", type == "ta_oa") |> pull(value) / publisher_summary_df |> filter(cr_year == "2022", publisher == "Elsevier") |> pull(value) |> sum()*100`%).

@@

Figure \ref{fig:subject_panel} presents the relative growth of hybrid open access by subject area between 2018 and 2022. In particular, Social Sciences including Arts and Humanities journals accounted for the strongest growth in the five-years period from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Social Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(value) |> sum()`), followed by the Life Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Life Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Life Sciences") |> pull(value) |> sum()`) , Health Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Health Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Health Sciences") |> pull(value) |> sum()`) and Physical Sciences from `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(prop) |> sum() * 100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2018", top_level == "Physical Sciences") |> pull(value) |> sum()`) to `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(prop) |> sum() *100`% (n = `r subject_trend_plot$data |> filter(cr_year == "2022", top_level == "Physical Sciences") |> pull(value) |> sum()`). 
<!-- In particular, journals in the subject areas recorded large high rate (see supplement)-->
Growth in Social Sciences category can be largely attributed to transformative agreements. 
In 2022, two-thirds of open access articles (`r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(prop) * 100`%, n = `r subject_ta_trend_plot$data |> filter(cr_year == "2022", top_level == "Social Sciences") |> pull(ta_oa_frac)`) were published by the first authors affiliated with participating institutions (see \ref{fig:subject_panel}B).

```{r subject_panel, fig.asp=0.8, fig.cap="Development of open access in hybrid journals in transformative agreements between 2018 and 2022 by ASJC subject area. The blue areas represent open access through transformative agreements, the grey areas depict open access articles where no link to an agreement could be established. (A) Proportion of open access articles in hybrid journals per year and subject area. (B) Percentage of hybrid open access via agreements per year and subject area Boxplots. (C) show individual open access uptake rates by individual hybrid journals and open access funding per publication year and subject area. The individual outliers are not shown. Note that data on transformative agreements ending before June 2021 were not available for this study."}

@@

Between 2018 and 2022, high-income countries almost exclusively dominated hybrid open access publishing through transformative agreements. 
To discern socio-economic differences, these countries were grouped according to their membership in the Organisation for Economic Co-operation and Development (OECD).
Overall, first authors affiliated with institutions from OECD member countries published `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa)` open access articles in hybrid journals, representing `r oecd_stats_global |> filter(cat == "OECD") |> pull(oa) / n_oa * 100`% of the investigated open access articles.
This disparity between OECD nations and other countries becomes even more evident when considering open access through transformative agreements, as  `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa)` of `r n_ta_oa`, or `r oecd_stats_global |> filter(cat == "OECD") |> pull(ta_oa) / n_ta_oa*100`% of open access articles were associated with such agreements.
